---
title: "Exploring Tidyverse"
output:
  html_document:
    df_print: paged
---

Testing how this notebook works!

```{r}
plot(cars)
```

**Tidyverse approach:**

```{r}
library(tidyverse)
```

Data is in "tidy format" when each row presents a single observation and each column is a variable

**most useful functions in dplyr** 
mutate() to add columns. 
select() to select and order columns. 
filter() to filter rows based on a boolean.

Lets get some data to play with:

```{r}
library(dslabs)
data("murders")
colnames(murders)
```

Adding another variable with mutate()

```{r}
murders <- mutate(murders, rate = total / population * 100000)  # add a new column 
```

Filtering rows

```{r}
murders_filt<- filter(murders, rate <= 0.71) # subsetting rows with with filter
colnames(murders_filt)
```

Subsetting columns with select

```{r}
new_table <- select(murders, state, region, rate) # you can change the order of the columns too. 
head(new_table)
```

**using the pipe \|\> or %\>%**

```{r}
murders |> select(state, region, rate) |> filter(rate <= 0.71)
# or just save the new df 
murders2 <- murders |> select(state, region, rate) |> filter(rate <= 0.71)
```

note how df is piped to various functions in tandem.
